Dynamic Witnesses for Static Type Errors

### 2015-11-20 paper/pldi16-submission/semantics.tex
address some WRW comments

% WRW is still minorly worried about this theorem as phrased (cf. every
% inhabitable type). Perhaps jsut a footnote explaining why \alpha -> \beta
% is not considered here?  
\begin{thm}{\textbf{[Witness Generality]}}
\label{thm:soundness}
  For any function $f$, if \steptr{\trace}{\eapp{f}{\ehole{}}}{\emptysu}{\stuck}{\su},
  then for every type $t$ there exists $\hastype{v}{t}$ such that
  $\steps{\eapp{f}{v}}{\emptysu}{\stuck}{\su}$.
\end{thm}

@@

% WRW: Just to be sure, the type of "lambda x. x + 1" would seem
% to be box_1 -> box_2 here, while one might be tempted to get
% int -> int. Some extra prose text contrasting typeof() with
% "apply f to box to get a partial type" would likely help. 
\paragraph{The Type of an Expression} The (dynamic) type of an
expression $e$ is defined as a function $\typeof{e}$ defined as:
  \[
  \begin{array}{lcll}
    \typeof{n}   & \defeq & \tint & \\
    \typeof{b}   & \defeq & \tbool & \\
    \typeof{\efun{x}{e}} & \defeq & \tfun{\thole{i}}{\typeof{e}}, & \quad \text{$i$ is fresh} \\
    \typeof{e} & \defeq & \thole{i}, & \quad \text{$i$ is fresh} \\
  \end{array}
  \]

>>>>

\begin{thm}{\textbf{[Witness Generality]}}
\label{thm:soundness}
  For any function $f$, if \steptr{\trace}{\eapp{f}{\ehole{}}}{\emptysu}{\stuck}{\su},
  then for every inhabitable type\footnote{We exclude builtin functions that subvert the type system, \eg \texttt{Obj.magic}, and thus consider the type $\forall a b. \tfun{a}{b}$ to be uninhabitable.}  $t$ there exists $\hastype{v}{t}$ such that
  $\steps{\eapp{f}{v}}{\emptysu}{\stuck}{\su}$.
\end{thm}

@@

\paragraph{The Type of an Expression} The (dynamic) type of an
expression $e$ is defined as a function $\typeof{e}$ defined as:
  \[
  \begin{array}{lcll}
    \typeof{n}   & \defeq & \tint & \\
    \typeof{b}   & \defeq & \tbool & \\
    \typeof{\efun{x}{e}} & \defeq & \tfun{\thole{i}}{\typeof{e}}, & \quad \text{$i$ is fresh} \\
    \typeof{e} & \defeq & \thole{i}, & \quad \text{$i$ is fresh} \\
  \end{array}
  \]
%
Note that we approximate the type of expressions rather than
performing full type inference, for example
$\typeof{\efun{x}{\eplus{x}{1}}} = \tfun{\thole{}}{\thole{}}$,
instead of the usual $\tfun{\tint}{\tint}$.
